
```{r check-na-strings}
# different na string spellings
na_strings <- c("N/A", "NA", "n/a", "na", "None", "NULL", "", "-", "unknown")

# count of na strings in each character column
train_clean %>%
  select(where(is.character)) %>%
  summarise(across(everything(), ~ sum(. %in% na_strings, na.rm = TRUE))) %>%
  pivot_longer(everything(), names_to = "column", values_to = "count") %>%
  filter(count > 0)

# list of unique na strings in each character column
train_clean %>%
  select(where(is.character)) %>%
  summarise(across(everything(), ~ list(unique(.[. %in% c("N/A", "NA", "n/a", "na", "None", "none", "NULL", "null", "", " ", "-", "--", "?", "unknown", "Unknown")]))))  %>%
  pivot_longer(everything(), names_to = "column", values_to = "na_strings") %>%
  unnest(na_strings) %>%
  filter(lengths(na_strings) > 0)

# list of unique na strings in each character column of test data
test_clean %>%
  select(where(is.character)) %>%
  summarise(across(everything(), ~ list(unique(.[. %in% c("N/A", "NA", "n/a", "na", "None", "none", "NULL", "null", "", " ", "-", "--", "?", "unknown", "Unknown")]))))  %>%
  pivot_longer(everything(), names_to = "column", values_to = "na_strings") %>%
  unnest(na_strings) %>%
  filter(lengths(na_strings) > 0)

```

- identified that there are different na string spellings in the data
- need to handle these na strings in the data cleaning pipeline otherwise they will be treated as categorical values, negatively impacting how the columns are treated (chr vs num)


```{r correlation-checks-train-set}
# na's of each variable
colSums(is.na(train_clean))

# distribution of host_response_time
table(train_clean$host_response_time)

# check rows where host_response_time is NA
train_clean %>% filter(is.na(host_response_time)) %>% print(width = Inf)

# check columns where both host_name and host_response_rate are na
train_clean %>% filter(is.na(host_name) & is.na(host_response_rate)) %>% print(width = Inf)

# check correlation between host_about and host_response_rate
# 1. make both columns numeric (binary, 0 if a value is not contained, 1 if there is one or more value)
train_clean %>%
  mutate(
    host_about_numeric = ifelse(is.na(host_about), 0, 1),
    host_response_rate_numeric = ifelse(is.na(host_response_rate), 0, 1)
  ) %>%
  select(host_about_numeric, host_response_rate_numeric) %>%
  cor()

# 2. create host_about_binary column
train_clean %>%
  mutate(
    host_about_binary = ifelse(is.na(host_about), 0, 1)
  )
# check na's in host_about_binary
sum(is.na(train_clean$host_about_binary))

# check correlation between host_response_rate and host_total_listings_count
# 1. convert both to binary
train_clean %>%
  mutate(
    host_response_rate_binary = ifelse(is.na(host_response_rate), 0, 1),
    host_total_listings_count_binary = ifelse(is.na(host_total_listings_count), 0, 1)
  ) %>% 
  select(host_response_rate_binary, host_total_listings_count_binary) %>%
  cor()

# check correlation between host_response_rate and host_response_time
# a. convert both to binary
train_clean %>%
  mutate(
    host_response_rate_binary = ifelse(is.na(host_response_rate), 0, 1),
    host_response_time_binary = ifelse(is.na(host_response_time), 0, 1)
  ) %>%
  select(host_response_rate_binary, host_response_time_binary) %>%
  cor()

# b. hosts w/out response data are likely newer or less active
# impute na's for host_response_rate as a new factor level: "unknown"
train_clean %>%
  mutate(
    host_has_response_data = as.factor(ifelse(is.na(host_response_rate), 0, 1)),
    host_response_rate = replace_na(host_response_rate, 0),
    host_response_time = as.factor(ifelse(is.na(host_response_time), "unknown", as.character(host_response_time)))
  )

sum(is.na(train_clean$host_response_rate))
sum(is.na(train_clean$host_response_time))
sum(is.na(train_clean$host_has_response_data))

# 2a. distribution of guests_included
table(train_clean$guests_included)

# 2b. check rows where guests_included is NA
train_clean %>% filter(is.na(guests_included)) %>% print(width = Inf)

# 2c. make binary features for guests_included, beds, bedrooms, bathrooms
train_clean %>%
  mutate(
    guests_included_binary = ifelse(is.na(guests_included), 0, 1),
    beds_binary = ifelse(is.na(beds), 0, 1),
    bedrooms_binary = ifelse(is.na(bedrooms), 0, 1),
    bathrooms_binary = ifelse(is.na(bathrooms), 0, 1),
    accommodates_binary = ifelse(is.na(accommodates), 0, 1)
  ) %>% 
  select(guests_included_binary, beds_binary, bedrooms_binary, bathrooms_binary, accommodates_binary) %>%
  cor()

# 3a. correlation of pricing-fields
train_clean %>%
  mutate(
    extra_people_binary = ifelse(is.na(extra_people), 0, 1),
    security_deposit_binary = ifelse(is.na(security_deposit), 0, 1),
    cleaning_fee_binary = ifelse(is.na(cleaning_fee), 0, 1),
    weekly_price_binary = ifelse(is.na(weekly_price), 0, 1)
  ) %>%
  select(extra_people_binary, security_deposit_binary, cleaning_fee_binary, weekly_price_binary) %>%
  cor()

# 4a. correlation of review features
train_clean %>% 
  select(review_scores_rating, review_scores_accuracy, review_scores_cleanliness, review_scores_checkin, 
         review_scores_communication, review_scores_location, review_scores_value) %>%
  cor(use = "complete.obs")

# 4b. check whether reviews_per_month na's align perfectly with the review score na's
train_clean %>%
  mutate(
    rating_binary = ifelse(is.na(review_scores_rating), 0, 1),
    rpm_binary = ifelse(is.na(reviews_per_month), 0, 1)
  ) %>%
  select(rating_binary, rpm_binary) %>%
  cor()

# 4c. reviews_per_month na's align with rating na's - will impute with 0 in chunk 3
# (shown here as exploratory only; actual imputation happens in impute-values-from-above)
```

```{r correlation-checks-test-set}
# na's of each variable
colSums(is.na(test_clean))

# distribution of host_response_time
table(test_clean$host_response_time)

# check rows where host_response_time is NA
test_clean %>% filter(is.na(host_response_time)) %>% print(width = Inf)

# check columns where both host_name and host_response_rate are na
test_clean %>% filter(is.na(host_name) & is.na(host_response_rate)) %>% print(width = Inf)

# check correlation between host_about and host_response_rate
# 1. make both columns numeric (binary, 0 if a value is not contained, 1 if there is one or more value)
test_clean %>%
  mutate(
    host_about_numeric = ifelse(is.na(host_about), 0, 1),
    host_response_rate_numeric = ifelse(is.na(host_response_rate), 0, 1)
  ) %>%
  select(host_about_numeric, host_response_rate_numeric) %>%
  cor()

# 2. create host_about_binary column
test_clean %>%
  mutate(
    host_about_binary = ifelse(is.na(host_about), 0, 1)
  )
# check na's in host_about_binary
sum(is.na(test_clean$host_about_binary))

# check correlation between host_response_rate and host_total_listings_count
# 1. convert both to binary
test_clean %>%
  mutate(
    host_response_rate_binary = ifelse(is.na(host_response_rate), 0, 1),
    host_total_listings_count_binary = ifelse(is.na(host_total_listings_count), 0, 1)
  ) %>% 
  select(host_response_rate_binary, host_total_listings_count_binary) %>%
  cor()

# check correlation between host_response_rate and host_response_time
# a. convert both to binary
test_clean %>%
  mutate(
    host_response_rate_binary = ifelse(is.na(host_response_rate), 0, 1),
    host_response_time_binary = ifelse(is.na(host_response_time), 0, 1)
  ) %>%
  select(host_response_rate_binary, host_response_time_binary) %>%
  cor()

# b. hosts w/out response data are likely newer or less active
# impute na's for host_response_rate as a new factor level: "unknown"
test_clean %>%
  mutate(
    host_has_response_data = as.factor(ifelse(is.na(host_response_rate), 0, 1)),
    host_response_rate = replace_na(host_response_rate, 0),
    host_response_time = as.factor(ifelse(is.na(host_response_time), "unknown", as.character(host_response_time)))
  )

sum(is.na(test_clean$host_response_rate))
sum(is.na(test_clean$host_response_time))
sum(is.na(test_clean$host_has_response_data))

# 2a. distribution of guests_included
table(test_clean$guests_included)

# 2b. check rows where guests_included is NA
test_clean %>% filter(is.na(guests_included)) %>% print(width = Inf)

# 2c. make binary features for guests_included, beds, bedrooms, bathrooms
test_clean %>%
  mutate(
    guests_included_binary = ifelse(is.na(guests_included), 0, 1),
    beds_binary = ifelse(is.na(beds), 0, 1),
    bedrooms_binary = ifelse(is.na(bedrooms), 0, 1),
    bathrooms_binary = ifelse(is.na(bathrooms), 0, 1),
    accommodates_binary = ifelse(is.na(accommodates), 0, 1)
  ) %>% 
  select(guests_included_binary, beds_binary, bedrooms_binary, bathrooms_binary, accommodates_binary) %>%
  cor()

# 3a. correlation of pricing-fields
test_clean %>%
  mutate(
    extra_people_binary = ifelse(is.na(extra_people), 0, 1),
    security_deposit_binary = ifelse(is.na(security_deposit), 0, 1),
    cleaning_fee_binary = ifelse(is.na(cleaning_fee), 0, 1),
    weekly_price_binary = ifelse(is.na(weekly_price), 0, 1)
  ) %>%
  select(extra_people_binary, security_deposit_binary, cleaning_fee_binary, weekly_price_binary) %>%
  cor()

# 4a. correlation of review features
test_clean %>% 
  select(review_scores_rating, review_scores_accuracy, review_scores_cleanliness, review_scores_checkin, 
         review_scores_communication, review_scores_location, review_scores_value) %>%
  cor(use = "complete.obs")

# 4b. check whether reviews_per_month na's align perfectly with the review score na's
test_clean %>%
  mutate(
    rating_binary = ifelse(is.na(review_scores_rating), 0, 1),
    rpm_binary = ifelse(is.na(reviews_per_month), 0, 1)
  ) %>%
  select(rating_binary, rpm_binary) %>%
  cor()

# 4c. reviews_per_month na's align with rating na's - will impute with 0 in chunk 3
# (shown here as exploratory only; actual imputation happens in impute-values-from-above)
```

```{r host-response-data-eda}
# NOTE: must run before select-features in chunk 4 (host_has_response_data is dropped there)

train_clean %>%
  group_by(host_has_response_data) %>%
  summarise(pct_high_booking = mean(as.integer(high_booking) - 1), n = n())

# does it correlate with city or other structural vars?
train_clean %>%
  group_by(host_has_response_data) %>%
  summarise(
    pct_high_booking = mean(as.integer(high_booking) - 1),
    median_price     = median(price, na.rm = TRUE),
    median_reviews   = median(reviews_per_month, na.rm = TRUE),
    n = n()
  )

# does response data missingness align with other missing fields?
train_clean %>%
  mutate(no_response_data = host_has_response_data == 0) %>%
  summarise(across(where(is.numeric), ~ mean(is.na(.) & no_response_data == TRUE)))

```

**1:**

- added new binary feature: host_about_binary
- added new binary feature: host_has_response_data
- added new categorical feature: host_response_time (with "unknown" as a level)

**2:**

- use accomodates as a proxy for other features (beds, bedrooms, bathrooms, guests_included)
    - accommodates being fully populated while beds/bedrooms have independent NAs means you can derive reasonable imputations like beds = ceiling(accommodates / 2) or bedrooms = floor(accommodates / 2) for missing rows rather than just using the global median
        - guests_included + bathrooms; Impute together; bathrooms→median, guests_included→1
            - has_property_details; shared binary flag bc essentially the same rows are missing (r = 0.995); capturing basic structural property details
        - beds; Impute via accommodates (e.g. accommodates / 2) or median
        - bedrooms; Impute via accommodates or median
        
**3:**

- extra_people + security_deposit + cleaning_fee (r = 0.68 - 0.8)
    - all optional fee fields that engaged/professional hosts tend to configure together
    - impute all three with 0
- weekly_price is independent (r = 0.16–0.27)

**4:**

- looked into correlation of all the review features
- all 7 review scores are nearly identical (r = 0.94–0.98) 
    - leading to multicollinearity issues, keeping all would add noise and redundancy
- between reviews_per_month and review_rating_binary, r = 0.967
    - some listings have a reviews_per_month but no individual score breakdowns
- given the extreme multicollinearity finding (r > 0.94), keep only review_scores_rating + reviews_per_month
