---
title: "eda: popincourt spatial patterns"
format: html
---

```{r library-setup}
#| include: false
library(tidyverse)
library(tidymodels)
library(plumber)
library(tidytable)
library(geosphere)
library(ggplot2)
library(sf)
library(rnaturalearth)
library(ggforce)
library(osmdata)
library(maptiles)
library(ggnewscale)
```

## phase 1: setup & first look

```{r load-data}
#| include: false
# load data with tidytable
airbnb_popincourt <- fread("C:/Users/ramse/msba-courses/ds 810/airbnb-project/part-two/data/airbnb-project-paris-sampled-Popincourt.csv")

# clean data
airbnb_popincourt <- airbnb_popincourt %>%
  mutate(
    price = as.numeric(gsub("[$,]", "", price)),
    amenity_count = lengths(strsplit(as.character(amenities), ",")),
    high_booking = as.factor(high_booking)
  )

# check data types
glimpse(airbnb_popincourt)

# check dimensions
dim(airbnb_popincourt)

# check na values per column
colSums(is.na(airbnb_popincourt))

```

## data visualization to explore statial patterns

```{r map-plot}
#| eval: true
#| echo: false
# plot listings on a map
ggplot(airbnb_popincourt, aes(x = longitude, y = latitude, color = high_booking)) +
  geom_point(alpha = 0.5) +
  theme_minimal()
```

```{r density-maps}
#| layout-ncol: 2
#| fig-height: 10
#| fig-width: 10
#| echo: false
#| eval: true

# hexbin map: avg high_booking rate per hex (0 = low, 1 = high)
airbnb_popincourt %>%
  mutate(high_booking_num = as.numeric(as.character(high_booking))) %>%
  ggplot(aes(x = longitude, y = latitude, z = high_booking_num)) +
  stat_summary_hex(bins = 30, fun = mean) +
  scale_fill_viridis_c(option = "B", name = "Avg High Booking Rate") +
  labs(
    title = "High Booking Rate Across Popincourt",
    x = "Longitude", y = "Latitude"
  ) +
  theme_minimal()

# hexbin map: avg price per hex
airbnb_popincourt %>%
  ggplot(aes(x = longitude, y = latitude, z = price)) +
  stat_summary_hex(bins = 30, fun = mean) +
  scale_fill_viridis_c(option = "B", name = "Avg Price (EUR)") +
  labs(
    title = "Average Listing Price Across Popincourt",
    x = "Longitude", y = "Latitude"
  ) +
  theme_minimal()

```

```{r density-dual-overlay}
#| echo: false
#| eval: true

# z-score price and high_booking at listing level, then flag alignment per listing
df_aligned <- airbnb_popincourt %>%
  mutate(
    high_booking_num = as.numeric(as.character(high_booking)),
    price_z   = as.numeric(scale(price)),
    booking_z = as.numeric(scale(high_booking_num)),
    aligned   = as.numeric(abs(price_z - booking_z) < 1)
  )

# stat_summary_hex averages the per-listing alignment flag per hex bin
# fill = proportion of listings in each hex where price and booking rate are within 1 SD
ggplot(df_aligned, aes(x = longitude, y = latitude, z = aligned)) +
  stat_summary_hex(bins = 30, fun = mean) +
  scale_fill_gradient(
    low = "#2D1B69", high = "#FFB347",
    name = "Prop. Aligned\n(price ~ booking)"
  ) +
  labs(
    title = "Price-Booking Alignment Across Popincourt",
    subtitle = "Bright hexes: listings where price rank and booking rate rank move together (within 1 SD)",
    x = "Longitude", y = "Latitude"
  ) +
  theme_minimal()
```

```{r dual-overlay-v2}
#| echo: false
#| eval: true

# rebuild both hex summaries at the hex level (same bins = 30)
p_booking <- airbnb_popincourt %>%
  mutate(high_booking_num = as.numeric(as.character(high_booking))) %>%
  ggplot(aes(x = longitude, y = latitude, z = high_booking_num)) +
  stat_summary_hex(bins = 30, fun = mean)

p_price <- airbnb_popincourt %>%
  ggplot(aes(x = longitude, y = latitude, z = price)) +
  stat_summary_hex(bins = 30, fun = mean)

hex_booking <- ggplot_build(p_booking)$data[[1]] %>%
  transmute(x, y, booking_rate = value)

hex_price <- ggplot_build(p_price)$data[[1]] %>%
  transmute(x, y, avg_price = value)

# flag hexes where BOTH are in the top quartile (bright/yellow in both plots)
hex_combined <- inner_join(hex_booking, hex_price, by = c("x", "y")) %>%
  mutate(
    both_high = booking_rate >= quantile(booking_rate, 0.75, na.rm = TRUE) &
                avg_price    >= quantile(avg_price,    0.75, na.rm = TRUE)
  )

ggplot(hex_combined, aes(x = x, y = y, z = as.numeric(both_high))) +
  stat_summary_hex(bins = 30, fun = mean) +
  scale_fill_gradient(
    low = "#2D1B69", high = "#FFB347",
    name = "Overlap Score"
  ) +
  labs(
    title = "High Price & High Booking Rate Overlap",
    subtitle = "Bright hexes: top quartile in both avg price AND avg high-booking rate",
    x = "Longitude", y = "Latitude"
  ) +
  theme_minimal()
```

```{r dual-overlay-v3}
#| echo: false
#| eval: true
library(osmdata)

# 1. get bounding box for Popincourt
bb <- getbb("Popincourt, Paris, France")

# 2. pull the neighborhood boundary
boundary <- opq(bb) %>%
  add_osm_feature(key = "boundary", value = "administrative") %>%
  osmdata_sf()

# 3. pull the street network
streets <- opq(bb) %>%
  add_osm_feature(key = "highway") %>%
  osmdata_sf()

# compute both_high at listing level so stat_summary_hex bins from raw coordinates
df_both_high <- airbnb_popincourt %>%
  mutate(
    high_booking_num = as.numeric(as.character(high_booking)),
    both_high = as.numeric(
      price >= quantile(price, 0.75, na.rm = TRUE) & high_booking_num == 1
    )
  )

options(scipen = 999)

ggplot() +
  geom_sf(data = streets$osm_lines, color = "grey50", linewidth = 0.2, inherit.aes = FALSE) +
  stat_summary_hex(
    data = df_both_high,
    aes(x = longitude, y = latitude, z = both_high),
    bins = 30, fun = mean, alpha = 0.75
  ) +
  geom_sf(data = boundary$osm_multipolygons, fill = NA, color = "white", linewidth = 0.9, inherit.aes = FALSE) +
  scale_fill_gradient(
    low = "#2D1B69", high = "#FFB347",
    name = "Overlap Score"
  ) +
  coord_sf(
    xlim = range(airbnb_popincourt$longitude) + c(-0.002, 0.002),
    ylim = range(airbnb_popincourt$latitude)  + c(-0.002, 0.002)
  ) +
  labs(
    title = "High Price & High Booking Rate Overlap",
    subtitle = "Bright hexes: top quartile in both avg price AND avg high-booking rate",
    x = "Longitude", y = "Latitude"
  ) +
  theme_minimal() +
  theme(panel.background = element_rect(fill = "#1a1a2e", color = NA))


```

## GOLDEN MAP BELOW: dual density overlay with satellite background showing top quartile in both avg price AND avg high-booking rate

use this map below in the report/presentation to show the spatial eda we performed which highlights (the yellow hexes) the areas with high price and high booking rate.

why use this map:

- shows the spatial distribution of high price and high booking rate
  - important to understand where the high-value listings are located
  - helps identify areas that may be experiencing gentrification
  

```{r dual-overlay-v4}
#| echo: false
#| eval: true

library(maptiles)
library(ggnewscale)

options(scipen = 999)

# bounding box from listing extent
bbox_sf <- st_bbox(
  c(xmin = min(airbnb_popincourt$longitude) - 0.002,
    xmax = max(airbnb_popincourt$longitude) + 0.002,
    ymin = min(airbnb_popincourt$latitude)  - 0.002,
    ymax = max(airbnb_popincourt$latitude)  + 0.002),
  crs = st_crs(4326)
)

# fetch Esri satellite tiles (no API key required)
satellite_tiles <- get_tiles(bbox_sf, provider = "Esri.WorldImagery", zoom = 15)

# pull Popincourt boundary from OSM
boundary <- opq(as.numeric(bbox_sf)) %>%
  add_osm_feature(key = "boundary", value = "administrative") %>%
  osmdata_sf()

# listing-level both_high flag
df_both_high <- airbnb_popincourt %>%
  mutate(
    high_booking_num = as.numeric(as.character(high_booking)),
    both_high = as.numeric(
      price >= quantile(price, 0.75, na.rm = TRUE) & high_booking_num == 1
    )
  )

# convert satellite SpatRaster to data frame for ggplot2 (no tidyterra needed)
tile_df <- as.data.frame(satellite_tiles, xy = TRUE)
names(tile_df)[3:5] <- c("r", "g", "b")
tile_df$color <- rgb(tile_df$r / 255, tile_df$g / 255, tile_df$b / 255)

ggplot() +
  geom_raster(data = tile_df, aes(x = x, y = y, fill = color)) +
  scale_fill_identity() +
  new_scale_fill() +
  stat_summary_hex(
    data = df_both_high,
    aes(x = longitude, y = latitude, z = both_high),
    bins = 30, fun = mean, alpha = 0.75
  ) +
  geom_sf(data = boundary$osm_multipolygons, fill = NA, color = "white", linewidth = 0.9, inherit.aes = FALSE) +
  scale_fill_gradient(
    low = "#2D1B69", high = "#FFB347",
    name = "Overlap Score"
  ) +
  coord_sf(
    xlim = c(bbox_sf["xmin"], bbox_sf["xmax"]),
    ylim = c(bbox_sf["ymin"], bbox_sf["ymax"])
  ) +
  labs(
    title = "High Price & High Booking Rate Overlap",
    subtitle = "Bright hexes: top quartile in both avg price AND avg high-booking rate",
    x = "Longitude", y = "Latitude"
  ) +
  theme_minimal()
```
