
```{r library-setup}
#| output: false
# libraries
library(tidyverse)
library(tidymodels)
library(vetiver)
library(plumber)
library(tidytable)
library(bonsai)
library(lightgbm)
library(pins)
library(beepr)
library(GPfit)
library(lhs)
library(dials)
library(caret)
library(parallel)
library(vip)
library(gtools)

# set seed
set.seed(21)

```


```{r data-loading}
airbnb_pop <- fread("C:/Users/ramse/msba-courses/ds 810/airbnb-project/part-two/data/airbnb-project-paris-sampled-Popincourt.csv")

glimpse(airbnb_pop)
```

```{r check-na}
# check na's in each column
colSums(is.na(airbnb_pop))
```

```{r data-preprocessing-for-logistic-regression}
# use only these varialbles for modeling
airbnb_pop <- airbnb_pop %>% 
    select(c(
        id, date, high_booking, host_id, host_since, host_location, host_response_time, host_response_rate, host_is_superhost, street, latitude, longitude, property_type, room_type, accommodates, bathrooms, bedrooms, beds, bed_type, square_feet, price, weekly_price, monthly_price, guests_included, extra_people, minimum_nights, maximum_nights, review_scores_rating, review_scores_accuracy, review_scores_cleanliness, review_scores_checkin, review_scores_communication, review_scores_location, review_scores_value, reviews_per_month, host_listings_count, host_total_listings_count, amenities, security_deposit, cleaning_fee, instant_bookable, minimum_minimum_nights, maximum_minimum_nights, minimum_maximum_nights, maximum_maximum_nights, minimum_nights_avg_ntm, maximum_nights_avg_ntm, latitude_cleansed, longitude_cleansed
    ))

glimpse(airbnb_pop)

# clean variable types
airbnb_pop <- airbnb_pop %>% 
    mutate(
        id = as.numeric(id),
        high_booking = as.factor(high_booking),
        date = as.Date(date),
        host_since = as.Date(host_since),
        host_tenure_days = as.numeric(date - host_since),
        listing_month = as.factor(format(date, "%m")),
        host_country = trimws(sub(".*,", "", host_location)),
        host_country = dplyr::recode(host_country,
            # ISO-2 country codes
            "FR" = "France",
            "DE" = "Germany",
            "GB" = "United Kingdom",
            "BE" = "Belgium",
            "CH" = "Switzerland",
            "ES" = "Spain",
            "IE" = "Ireland",
            "IN" = "India",
            "JP" = "Japan",
            "HK" = "Hong Kong",
            "DK" = "Denmark",
            "KG" = "Kyrgyzstan",
            "CM" = "Cameroon",
            "PT" = "Portugal",
            "QA" = "Qatar",
            "RS" = "Serbia",
            "AL" = "Albania",
            "FI" = "Finland",
            # US state codes (from "City, ST" entries)
            "NY" = "United States",
            "CA" = "United States",
            "TX" = "United States",
            # Non-standard / city-only entries
            "LONDON" = "United Kingdom",
            "London" = "United Kingdom",
            "quebec" = "Canada"
        ),
        host_country = as.factor(host_country),
        host_response_time = as.factor(ifelse(host_response_time == "N/A", NA, as.character(host_response_time))),
        host_response_rate = as.numeric(gsub("%", "", host_response_rate)),
        host_is_superhost = as.factor(host_is_superhost),
        street_name = as.factor(sub(",.*", "", street)),
        property_type = as.factor(property_type),
        room_type = as.factor(room_type),
        bed_type = as.factor(bed_type),
        weekly_price = as.numeric(gsub("[$,]", "", weekly_price)),
        monthly_price = as.numeric(gsub("[$,]", "", monthly_price)),
        extra_people = as.numeric(gsub("[$,]", "", extra_people)),
        security_deposit = as.numeric(gsub("[$,]", "", security_deposit)),
        cleaning_fee = as.numeric(gsub("[$,]", "", cleaning_fee)),
        instant_bookable = as.factor(instant_bookable),
        price = as.numeric(gsub("[$,]", "", price)),
        price_per_person = price / accommodates,
        amenity_count = ifelse(is.na(amenities), 0, lengths(strsplit(as.character(amenities), ","))),
        reviews_x_rating = reviews_per_month * review_scores_rating,
        price_x_bedrooms = price / (bedrooms + 1),
        )

# ensure these are all the correct data types
glimpse(airbnb_pop)
```

```{r select-features}
# select features
airbnb_features <- airbnb_pop %>%
    select(c(
        # target
        high_booking,
        # host features
        host_tenure_days,
        host_country,
        host_response_time,
        host_response_rate,
        host_is_superhost,
        host_listings_count,
        host_total_listings_count,
        # listing features
        listing_month,
        street_name,
        property_type,
        room_type,
        accommodates,
        bathrooms,
        bedrooms,
        beds,
        bed_type,
        latitude_cleansed,
        longitude_cleansed,
        # pricing features
        price,
        price_per_person,
        price_x_bedrooms,
        extra_people,
        # booking policies
        guests_included,
        minimum_nights,
        maximum_nights,
        instant_bookable,
        # review features
        review_scores_rating,
        review_scores_accuracy,
        review_scores_cleanliness,
        review_scores_checkin,
        review_scores_communication,
        review_scores_location,
        review_scores_value,
        reviews_per_month,
        reviews_x_rating,
        # amenity features
        amenity_count
    ))
```

```{r check-na-features}
# check for missing values
colSums(is.na(airbnb_features))
```

```{r train-test-split}
# train/test split stratified on target
airbnb_split <- initial_split(airbnb_features, prop = 0.8, strata = high_booking)

airbnb_train <- training(airbnb_split)
airbnb_test <- testing(airbnb_split)
```

```{r recipe}
airbnb_recipe <- recipe(high_booking ~ ., data = airbnb_train) %>%

    # feature engineering
    # flag listings with no reviews before imputing scores
    step_mutate(has_reviews = as.factor(ifelse(is.na(reviews_per_month), 0, 1))) %>%

    # imputation
    # review scores: missing = no reviews yet, impute with 0
    step_impute_median(
        review_scores_rating, review_scores_accuracy, review_scores_cleanliness,
        review_scores_checkin, review_scores_communication, review_scores_location,
        review_scores_value, reviews_per_month, reviews_x_rating
    ) %>%
    # street_name: spatial nearest-neighbor using lat/long
    step_impute_knn(
        street_name,
        impute_with = imp_vars(latitude_cleansed, longitude_cleansed),
        neighbors = 1
    ) %>%
    # remaining numeric: median
    step_impute_median(all_numeric_predictors()) %>%
    # remaining nominal: mode
    step_impute_mode(all_nominal_predictors(), -all_outcomes()) %>%

    # handle high-cardinality factors
    step_other(street_name, threshold = 0.01) %>%
    step_other(host_country, threshold = 0.01) %>%

    # encoding
    step_dummy(all_nominal_predictors()) %>%

    # cleanup
    step_zv(all_predictors()) %>%
    step_normalize(all_numeric_predictors())
```

```{r logistic-regression-v1}
# model spec
logistic_spec <- logistic_reg() %>%
    set_engine("glm") %>%
    set_mode("classification")

# workflow
airbnb_workflow <- workflow() %>%
    add_recipe(airbnb_recipe) %>%
    add_model(logistic_spec)

# fit
logistic_fit <- airbnb_workflow %>%
    fit(data = airbnb_train)

tidy(logistic_fit) %>% 
    mutate(signif = stars.pval(p.value)) %>%
    print(n = Inf)

# filter for only significant predictors
tidy(logistic_fit) %>% 
    mutate(signif = stars.pval(p.value)) %>%
    filter(signif %in% c("*", "**", "***")) %>%
    print(n = Inf, width = Inf)
```



## Logistic Regression: Coefficient Interpretation

The model identified **57 statistically significant predictors** (p < 0.05) of high booking activity. The coefficients below are on the log-odds scale, where positive values increase the probability of high booking and negative values decrease it.

### Pricing

All three price variables `price`, `price_per_person`, and `price_x_bedrooms` are strongly negative, with `price_per_person` carrying the largest effect (−1.09). This is the single clearest signal in the model: higher per-guest cost reduces booking likelihood. Listings with lower prices relative to their capacity and bedroom count consistently outperform.

### Stay Requirements

Both `minimum_nights` (0.226) and `maximum_nights` (0.194) are positive and among the strongest predictors. This likely reflects that high-booking listings tend to attract longer-term stays, guests who commit to a stay are less likely to cancel and more likely to rebook.

### Host Behavior

`host_response_rate` (0.283) and `host_response_time_within.an.hour` (0.389) are both positively associated with high booking. Hosts who respond quickly and consistently drive more conversions. Interestingly, `host_is_superhost` is negative (−0.125), which may reflect that superhosts price at a premium, reducing booking volume despite higher quality.

### Reviews

`reviews_per_month` (−0.816) and `review_scores_rating` (−0.810) are both strongly negative, seemingly counterintuitive. The likely explanation is that high-booking listings tend to be newer or more available, and therefore have fewer accumulated reviews. `has_reviews_X1` being negative (−0.124) supports this: listings *without* any reviews have a higher predicted booking rate, consistent with newer listings receiving more initial traffic. `review_scores_value` is the exception, with a positive effect (0.538), confirming that perceived value matters more to guests than overall rating.

### Location & Property Type

`street_name_Paris` is a strong positive predictor (0.366), indicating centrally-addressed listings outperform. Among property types, `Entire.apartment` is positive (0.101) while `Entire.rental.unit` (−0.326) and `Entire.loft` (−0.099) reduce booking probability. Guests prefer traditional apartment-style listings. Shared and private room types are all negative relative to the baseline, consistent with guests preferring full properties.

### Seasonality

Listing months May through August are all negative relative to the reference month, with July showing the largest drop (−0.124). This may reflect increased supply during peak summer months, making it harder for individual listings to stand out.

### Amenities

`amenity_count` is negative (−0.196), which is unexpected. A likely explanation is that listings with many amenities are priced higher, and the negative price effect dominates. It may also reflect that heavily-listed amenities signal a more commercialized listing that guests are less drawn to.

```{r roc-auc}
# predictions on test set
logistic_test_results <- airbnb_test %>%
    bind_cols(
        predict(logistic_fit, airbnb_test, type = "prob"),
        predict(logistic_fit, airbnb_test, type = "class")
    )

# ROC curve and AUC
roc_curve_data <- roc_curve(
    data = logistic_test_results,
    truth = high_booking,
    .pred_1,
    event_level = "second"
)
roc_auc_val <- roc_auc(
    data = logistic_test_results,
    truth = high_booking,
    .pred_1,
    event_level = "second"
)

# print results
print(roc_curve_data)
print(roc_auc_val)
```

## Model Performance

The logistic regression model achieved an AUC of **0.859** on the held-out test set, indicating strong discriminative ability between high and non-high booking listings. An AUC in the 0.80–0.90 range is generally considered good, meaning the model correctly ranks a randomly chosen high-booking listing above a randomly chosen non-high-booking listing approximately 86% of the time.

This is a notable result for a baseline logistic regression with no hyperparameter tuning, suggesting the engineered features, particularly pricing variables, stay requirements, and review activity contain substantial predictive signal. The model provides a solid foundation for comparison against more complex approaches such as regularized regression or tree-based models.

